---
---

# Overview {.chapter}

The **rmdmux** package is an **R** package, which extends [rmarkdown](https://github.com/rstudio/rmarkdown).
It offers two additional output formats (HTML templates):

- `html_rmdmux_paged`: A beautiful **PDF** report
- `html_rmdmux_plain`: A plain **HTML** file, which fulfills the [WAI requirements](https://www.w3.org/WAI)
  and is therefore fully accessible by assistive reading devices like screen readers
  
With **rmdmux** you can create **a single RMD file** (see [rmarkdown](https://github.com/rstudio/rmarkdown)) and export it as PDF and
as accessible HTML file.

> Documents should be beautiful and accessible! 

The **rmdmux** package is a minor extension/variation of the wonderful
**[pagedown](https://pagedown.rbind.io/)** package and therefore
uses the **[pagedjs](https://pagedjs.org)** Javascript library in order to
create paginated HTML files.

## Install {#sec-install}

Please install the following R packages from **github**:

- [funky](https://github.com/a-maldet/funky)
- [composerr](https://github.com/a-maldet/composerr)
- [styledTables](https://github.com/R-package/styledtable) (v0.2.0 - deprecated, will hopefully be replaced by **styledtable** till May 2023)
- [xlsxcollection](https://github.com/a-maldet/xlsxcollection) (Optional: Used for Excel export)
- [rmdmux](https://github.com/a-maldet/rmdmux)

```{r eval = FALSE, echo = TRUE}
devtools::install_github('a-maldet/funky', build_opts = NULL)
devtools::install_github('a-maldet/composerr', build_opts = NULL)
devtools::install_github('R-package/styledTables', ref = "v0.2.0", build_opts = NULL)
devtools::install_github('a-maldet/rmdmux', build_opts = NULL)

# Optional (needed for Excel export)
devtools::install_github('a-maldet/xlsxcollection', build_opts = NULL)
```

## Getting started {.newpage}

In the following section we will have a look at the build-in example `ex1`.
We can copy the entire example into our current working directory by 
executing the following code:

```{r eval = FALSE, echo = TRUE}
# copy `ex1/` to current working directory
rmdmux::create_example_1(".")
```

The code above copies the following files into your working directory:

- `ex1_main.Rmd`: The main file containing the **yaml header** and linking the **distributed Rmd files** `ex1_child1.Rmd` and `ex1_child2.Rmd`
- `ex1_child1.Rmd` and `ex1_child2.Rmd`: Distributed chapters of the report
- `_bookdown.yml`: A **yaml** file configuring the used language translations for
  important report elements like **Figure**, **Table** etc. - see
  [https://bookdown.org/yihui/bookdown/internationalization.html](https://bookdown.org/yihui/bookdown/internationalization.html).
  If the report is rendered and the `_bookdown.yml` file is located in the same
  folder as the **Rmd** file, then the definitions in `_bookdown.yml` will
  automatically be included.
- `ex1_paged.sass`: An additional stylesheet, holding additional styling definitions
  for the reports paginated output format `html_rmdmux_paged`. This stylesheet
  is included in the **yaml header**.
- `ex1_plain.sass`: An additional stylesheet, holding additional styling definitions
  for the reports plain html output format `html_rmdmux_plain`. This stylesheet
  is included in the **yaml header**.
- `ex1_mytable.sass`: An additional stylesheet, which will be used in order
  to define a custom styling for table \@ref(tab:usa-arrests-customized).
  This stylesheet is included in the chunk where the table is created.
- `ex1_cover.png`: An optional image, which will be placed in the center of the reports cover page.
  This image is included by setting the `title-page:image` option in the **yaml header**.
- `ex1_logo.png`: An optional image, which will be placed in the corner of the reports cover page.
  This image is included by setting the `title-page:logo` option in the **yaml header**.
- `ex1_qrcode.png`: An optional image, which contains a qr-code and is used as
  additional url visualisation on the imprint page (2nd page) of the report.
  This image is included by setting the `imprint:url-qrcode` option in the **yaml header**.

### Rendering the report

As we can see in `ex1/ex1_run.R`, there are several ways to render the report:

- `rmarkdown::render("ex1_main.Rmd", output_format = "rmdmux::html_rmdmux_plain")`: 
  Calling `rmarkdown::render()` is the usual way for rendering **Rmd** files.
  Be aware that while this will work fine for the plain HTML output (`html_rmdmux_plain`),
  it won't for the paginated output (`html_rmdmux_paged`), since we will need a
  local HTTP server server, serving all required Javascript files.
- `rmdmux::print_chrome("ex1_main.Rmd", output_format = "rmdmux::html_rmdmux_paged")`:
  This call prints the paginated report to a PDF file. Be aware that
  you need to have **Chrome** installed.
- `rmdmux::inf_mr("ex1_main.Rmd", output_format = "rmdmux::html_rmdmux_paged")`:
  This call is very useful, when you are working on report, since it
  will create a local **HTTP server**, serving the rendered HTML document.
  You just need to open the url
  (something like `127.0.0.1:4321/ex1_main.html`) printed in the console output
  in **Google Chrome** and you will see the current state of your report.
  If you change one of your RMD files the report will automatically be 
  rerendered and you just need to refresh your browser.
  In order to quit the HTTP server, just call `rmdmux::daemon_stop()`

### Supported features

- Almost all features of [pagedown](https://pagedown.rbind.io/)
  (except **running headers** and **front-cover** and **back-cover images**):
    - table of contents
      (see [bookdown/#table-of-contents](https://bookdown.org/yihui/rmarkdown/html-document.html#table-  of-contents))
    - list of figures (see [pagedown/#LOT](https://pagedown.rbind.io/#LOT))
    - list of tables (see [pagedown/#LOT](https://pagedown.rbind.io/#LOT))
    - customized labels (language support) for report elements
      (see [pagedown/#internationalization](https://pagedown.rbind.io/#internationalization))
    - footnotes can be included in the page footer of the corresponding pages
      (see [pagedown/#footnotes](https://pagedown.rbind.io/#footnotes))
    - referencing figures, tables, sections etc. with counter and page number
      (see [pagedown/#page-references](https://pagedown.rbind.io/#page-references))
      (Cation: The label for the page number references is defined in **css** 
      and not in **\_bookdown.yaml**)
    - additional styling for the plain and the paginated report version by using
      **sass** or **css** files
      (see [pagedown/#the-css-overriding-mechanism](https://pagedown.rbind.io/#the-css-overridin    g-mechanism))
    - mathematical formulas in LaTeX
      (see [pagedown/#mathjax](https://pagedown.rbind.io/#mathjax))
    - automated creation of bibliography
      (see [pagedown/#bibliography](https://pagedown.rbind.io/#bibliography))
    - automated creation of a list of abbreviations
      (see [pagedown/#list-of-abbreviations](https://pagedown.rbind.io/#list-of-abbreviations))
- title page with
    - logo: optional small image in the upper right corner of the title page
    - image: optional big image in the center of the title page
    - footer: optional title page footer with left, center and right page footer
      area
- imprint page with:
    - report title field
    - edition field: Optional field holding the print edition
    - date field: Publishing date
    - place field: Optional field holding the place of pulication
    - publisher field: Optional field holding the name publisher name
    - author field: Field listing all authors
    - editor: Optional field listing all editors
    - address: Optional field holding the publisher address
    - phone number field: Optional field with phone number
    - email field: Optional field with email address
    - url field: Optional field with a publisher or report url
    - qr-code: Optional qr-code image
    - credits field: Optional credits for images etc.
    - remarks area: Optional remarks at the end of the imprint page
- fully accessible tables by using
  [styledTables](https://github.com/R-package/styledtable):
    - beautiful tables, which fulfill the
      [WAI requirements](https://www.w3.org/WAI) and are accessible by
      screen readers.
    - tables can optionally be included in the list of tables and be referenced by
      `\@ref(tab:CHUNKNAME)`
    - tables can have a table caption and table footer
    - tables can contain multiple header rows and multiple row header columns
    - tables can contain subheader rows splitting the table into muliple areas
    - tables can be suppressed/included depending on the rendered output format
    - the appearance of the tables can be modified by supplying custom
      **sass** or **css** files
- splitting up the **rmd** file into multiple files

### Not supported features {.newpage}

The following features are supported by [pagedown](https://pagedown.rbind.io/),
but not by **rmdmux**:

- **front-cover** and **back-cover**
  (see [pagedown/#covers](https://pagedown.rbind.io/#covers)): This feature
  will most likely not be included anytime soon.
- **running headers**
  (see [pagedown/#custom-running-headers](https://pagedown.rbind.io/#custom-running-headers)):
  This feature will be included as soon as possible.


### Create your own styles

Have a look at the following stylesheets:

- [rmdmux/inst/stylesheets/rmdmux_paged_default.sass](https://github.com/a-maldet/rmdmux/trees/master/inst/stylesheets/rmdmux_paged_default.sass) Default stylesheet for **paged HTML output**
- [rmdmux/inst/stylesheets/rmdmux_plain_default.sass](https://github.com/a-maldet/rmdmux/trees/master/inst/stylesheets/rmdmux_plain_default.sass): Default stylesheet for **plain HTML output**

You may want to modify them and include the new stylesheets in the
`html_rmdmux_paged:css` and `html_rmdmux_plain:css` options of the reports **YAML** header.
