---
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.topcaption = TRUE,
  tab.add_dummy_table = TRUE,
  message = FALSE
)
```

# Supported features by **rmdmux** {.chapter}

## Distributed RMD files

For longer reports it can be useful to distribute the
code among multiple smaller RMD files. In order to include an RMD file
(child file) in another RMD file, one must include an empty chunk
with setting the child files name in the chunk option `child="FILENAME_CHILD.Rmd"`.

## Footnotes

For `html_rmdmux_paged` we can have paginated
footnotes^[For a description on footnotes see [pagedown/#footnotes](https://pagedown.rbind.io/#footnotes).]. This is
achieved by setting `pandoc_args: ["-M", "paged-footnotes: true"]` in
the YAML-Header.
For `html_rmdmux_plain` we cannot use paginated footnotes.
Because of this, we have cannot just use the general option 
`paged-footnotes: true`, since this is option is not valid for `html_rmdmux_plain`.

## Referincing tables, figures, sections

In section \@ref(sec-install) we saw, how **rmdmux** could be installed.
We will now create the accessible tables \@ref(tab:usa-arrests), 
\@ref(tab:usa-arrests-customized) and the figure \@ref(fig:phones) with
a conditional table (`html_rmdmux_plain` only).

## Creating accessible tables

The **[styledTables](https://github.com/R-package/styledtable)** package allows
the creation of accessible **HTML tables**. When creating the
**plain HTML** using the `html_rmdmux_plain` output format, the resulting
HTML tables will fulfill the 
[WAI requirements](https://www.w3.org/WAI) and therefore contain
all required features in order to be
fully accessible by screen reading devices, which are often used by visually impaired
persons.
Therefore, even very complex tables can be made accessible.
Supported features:

- **complex column headers** with multiple header levels and merged cells
- **complex row headers** with multiple header levels and merged cells
- **multiple subheadings** partitioning the table into sub tables

### Include styledTables in the **list of tables**

Unfortunately, **pagedown** currently only supports **Markdown** tables.
This means that an **HTML table** will not be included in the generated
**list of tables** (see the following [pagedown issue](https://github.com/rstudio/pagedown/issues/191)).
As a workaround, we can set the chunk option `tab.add_dummy_table = TRUE`
at the beginning of the **main RMD** file.
This causes the **styledTables** package to automatically add
**dummy tables** before each printed **styledtable**.
These dummy tables are only needed for the creation of the **list of tables** and
will automatically be stripped by an included pandoc-filter after this step.

### Create a simple styledTable {.newpage}

The following code prints a simple **styledTable**.
The **table caption** and the **table footer** are defined in the
chunk options. The **table reference** is defined by the **chunk name**
(do not use underscores or white spaces in chunk names and be sure that
all chunk names in the entire report are unique).
The table can be refernced as follows `\\@ref(tab:usa-arrests)` resulting in
the link \@ref(tab:usa-arrests) referencing the printed table.
These references also work for distributed **Rmd** files.

A table usually consists of the following elements:
 
- A **table caption**, which holds a table reference and describes the contents of
  the table. The table caption is defined by the chunk option `tab.cap`.
- An optional **table footer**, which consists of one or several text blocks
  containing additional information about the table. The table footer can be
  defined by the chunk option `tab.footer`.
- **Column header rows**: These rows are usually at the top of the table and
  they hold the headings of the data columns. The function
  `set_html_colheader()` can be used in order to tell the styledTables package
  which rows are holding the column headings.
- **Row header columns**: These columns are usually at the left side of the table
  and they hold the headings of the data rows. The function
  `set_html_rowheader()` can be used in order to tell the styledTables package
  which rows are holding the column headings.
- **Sub headings**: These rows are usually single merged rows, which split
  the data table into multiple small sub tables and each sub table
  contains only the data valid for the context given by the corresponding
  sub heading. The function `set_html_subheading()` can be used in order to tell
  the styledTables package which rows are sub heading rows.

```{r usa-arrests, tab.cap = 'Example-1 - A simple table', tab.footer = c('This is the table footer.', 'The footer may contain multiple rows.')}
# create styledTable object
st_usarrests <- df_usarrests %>%
  styled_table(keep_header = TRUE) %>%
  set_html_colheader(row_id = 1) %>%
  set_html_rowheader(col_id = 1)

# apply default theme the the styledTable
st_usarrests %>%
  apply_html_theme1() %>%
  xlsx_add_st
```

The function `apply_html_theme1()` applies the table styling defined in the
stylesheet in [styledTables/inst/sass_files/styledtable_theme1.sass](https://github.com/R-package/styledtable/trees/master/inst/sass_files/styledtable_theme1.sass).

### Apply your own table theme

In order to use your own table styling, you may want to have a look at the
default stylesheet
[styledTables/inst/sass_files/styledtable_theme1.sass](https://github.com/R-package/styledtable/trees/master/inst/sass_files/styledtable_theme1.sass)
and create your own theme from it.

In the following example we

- create a stylesheet `ex1_mytable.sass` defining the css styling for
  tables of HTML class `mytable`
- transpile the stylesheet to `ex1_mytable.min.css`
- add the created **css** file to the **html dependencies** of the styled table object.
- add the **HTML class** `ex1-mytable` to the styled table object.

```{r usa-arrests-customized, tab.cap = 'Example-2 - A table with custom theme'}
# transpile to css
sass::sass(
  sass::sass_file("./ex1_mytable.sass"),
  output = "ex1_mytable.min.css",
  options = sass::sass_options(output_style = "compressed"),
  cache = FALSE
) %>% invisible 

# add html dependency
st_usarrests_custom <- st_usarrests %>%
  add_html_stylesheet("ex1_mytable.min.css")

# add the html class "my-styledtable" and print the table
st_usarrests_custom %>%
  set_html_table_class("ex1-mytable") %>%
  xlsx_add_st
```

### Create a complex styledTable {.newpage}

```{r hair-eye-color, tab.cap = 'Example-3 - A more complex table', warnings = FALSE}
st_haireye_body <- df_haireye %>%
  styled_table

# Create the styled table header
st_haireye_header <- c( 
  "Hair", rep(c("Male", "Female"), each = 4),
  "", rep(c("Brown", "Blue", "Hazel", "Green"), times = 2)
) %>%
  styled_table(nrow = 2) %>%
  merge_cells(col_id = 1, row_id = 1:2) %>%
  merge_cells(col_id = 2:5, row_id = 1) %>%
  merge_cells(col_id = 6:9, row_id = 1)

# concatenate the styled table header and body,
# define col-/row-headers,
# apply `theme1` from `styledTables` package and then print the table
styled_table(
  st_haireye_header,
  st_haireye_body
) %>%
  set_html_colheader(row_id = 1:2) %>%
  set_html_rowheader(col_id = 1) %>%
  apply_html_theme1() %>%
  xlsx_add_st
```

### Conditional tables for plots {.newpage}

Plots are a great way to visualize data, but are not accessible to visually
impaired people.
Therefore, it is necessary to include a text or
an accessible table describing the displayed data.
With **styledTables** it is possible to append tables to figures only
for the **plain HTML output**.
This can be achieved by using the chunk option `tab.print = rmdmux::rmdmux_is_plain()`,
which suppresses the printing of the additional styled table, if the
current output format is not `html_rmdmux_plain`.

```{r, eval = FALSE, fig.height = 2, fig.cap="Number of telephones in various regions of the world (in thousands)", tab.print = rmdmux::rmdmux_is_plain()}
df_phones %>%
  pivot_longer(cols = 2:8, names_to = "region", values_to = "phones") %>%
  {
    ggplot(.) +
      geom_line(aes(x = year, y = phones, color = region))
  }

styled_table(df_phones, keep_header = TRUE) %>%
  styled_table %>%
  set_html_colheader(row_id = 1) %>%
  set_html_rowheader(col_id = 1) %>%
  apply_html_theme1()
```



```{r phones, echo = FALSE, fig.height = 2, fig.cap="Number of telephones in various regions of the world (in thousands)", tab.print = rmdmux::rmdmux_is_plain(), fig.footer = "In the `plain` export this plot is extended by an additional table holding the data. Note that, the chunk argument `fig.footer` is introduced in `rmdmux` for adding additional footer texts to figures."}
df_phones %>%
  pivot_longer(cols = 2:8, names_to = "region", values_to = "phones") %>%
  {
    ggplot(.) +
      geom_line(aes(x = year, y = phones, color = region))
  }

styled_table(df_phones, keep_header = TRUE) %>%
  styled_table %>%
  set_html_colheader(row_id = 1) %>%
  set_html_rowheader(col_id = 1) %>%
  apply_html_theme1()
```

### Store the tables in an Excel file

The package [xlsxcollection](www.github.com/a-maldet/xlsxcollection) allows you to addionally store the created
**styledTable** objects in an Excel file.
In this example we used the function `xlsx_add_st`, which saves the
table as an RDS file (this is a side effect) to the folder `xlsx_tables` and 
then simply returns the styledTable object, which causes **knitr** to print
the table.

Note: If you removed the call `%>% xlsx_add_st()` then the table would be
rendered as usual, but not stored in the folder `xlsx_tables`.
Hence, if you do not want to use the Excel export at all, you can
simply delete the chunk `excel-export` in `ex1_main.Rmd` and all calls
`%>% xlsx_add_st()` in the following chunks.

